---
title: "Topics in reproducible social science"
author: "Mats Ahrenshop"
date: "20 May 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reproducibility and research transparency

1. Ethical research
- IRB
- Fraud \& unintentional bias

2. Pre-registration and pre-analysis plans
- Purpose: publication bias \& researcher degrees freedom
- [EGAP](http://egap.org/design-registrations)
- [PAP example](https://github.com/mahrenshop/experimental_methods2020/blob/master/lab%20materials/week%204/christensen%20et%20al%20PAP.pdf)

3. Replication and reproducibility
- Reproducible code
- Data sharing
- [R Markdown](https://r4ds.had.co.nz/communicate-intro.html)
- [github](https://happygitwithr.com/)

# Pre-Analysis Plans

## Research hypotheses
1. Causal parameter (estimand) + threats to inference via exclusion restriction and non-interference
2. Sign and magnitude: prior beliefs about parameter
3. Heterogeneous treatment effects: specification of important subgroups


## Treatment description

- Context
- Treatment administration
- Information environment

## Inclusion of subjects
- Criteria for inclusion / eligibility for random assignment
- Population and sampling method
- Sample restrictions

## Random assignment

- Assignment procedure: via random number seed and automated procedure
- Handling of bad randomizations
- Simple (complete) random assignment vs blocked assignment vs clustered assignment
- Probabilities of assignment to conditions
- Flow diagram of allocation procedure (sampling, allocation, delivery of treatment, attrition)

## Soundness of randomization procedure

- Covariate balance: integrity of randomization procedure
- Magnitude of imbalance and statistical tests
- Randomization inference: imbalance magnitude expected by chance?
- Review random assignment procedure

## Outcome measures

- Justification for measurement instruments
- Validity
- Describe attrition
- Collection of outcome data
- Define what constitute non-compliance with treatment

## Data analysis

- Distribution of outcomes for each condition
- Avg. outcomes + sd for each condition
- ATE + SE
- Regression with and without covariates
- Randomization inference to test sharp H0 no effect
- Variance in outcomes, test for heterogeneity
- Address non-compliance, interference, and attrition

## Ethics
- IRB approval for research involving human subjects
- Discussion of
  * Harmful interventions and risk to subjects (minimal risk; mitigate losses; induce positive treatments)
  * Obtaining informed consent and secure withdrawal from study: subjects understand information on risk; voluntarily enter into an agreement: actively consent; may withdraw from participation
  *Protection of anonymity and confidentiality


# DeclareDesign

Describes research design in code (declare design); simulate designs to understand their properties (diagnose design) (such as power, bias, RMSE).

Load packages

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(DeclareDesign)
library(fabricatr)
library(randomizr)
library(estimatr)
library(DesignLibrary)
```

Declare a design

```{r, warning=FALSE, message=FALSE}
set.seed(1234)

simple_design <-
  # M: model
  
  # a 100-unit population with an unobserved shock 'u'
  
  declare_population(N = 100, u = rnorm(N)) +
  
  # 2 potential outcomes, where Y_Z_0 is the unobserved shock 'u'
  # and Y_Z_1 is that plus treatment effect, let's say 0.25
  
  declare_potential_outcomes(Y_Z_0 = u, Y_Z_1 = Y_Z_0 + 0.25) +
  
  
  # I: inquiry
  
  declare_estimand(PATE = mean(Y_Z_1 - Y_Z_0)) +
  
  # D: data strategy
  
  # we randomly sample 50 of the 100 units in the population
  
  declare_sampling(n = 50) +
  
  # randomly assign half of the 50 to treatment and control
  
  declare_assignment(prob = 0.5) +
  
  # construct observed outcomes from potential outcomes depending on realized assignment Z
  declare_reveal(outcome_variables = Y, assignment_variables = Z) +
  
  # A: answer strategy
  
  declare_estimator(Y ~ Z, model = difference_in_means, estimand = "PATE")

simple_design
```


For diagnosing designs, we define a set of diagnosands (statistical properties of the design, such as power, bias, RMSE).


```{r, message=FALSE, warning=FALSE}
simple_design_diagnosands <-
  declare_diagnosands(select = c(bias, rmse, power))

# now simulate design again and again, and calculate diagnosands based on the simulation data
# simulations of the DGP, then calculate estimands, and estimates

simple_design_diagnosands <-
  diagnose_design(simple_design, diagnosands = simple_design_diagnosands, sims = 500)

simple_design_diagnosands
```


The package \texttt{DesignLibrary} has many designs stored already:

```{r, message=FALSE, warning=FALSE}
b_c_design <- block_cluster_two_arm_designer(N = 1000, N_blocks = 10)
b_c_design

diagnose_design(b_c_design)
```



